Paper template for {ICPhS} 2023 Prague

### 2024-04-25 manuscript/main.tex

Interestingly, Medieval Spanish, similarly to Modern Catalan,
distinguished between /s/ and /z/ in intervocalic position within a
word, although these sounds merged through a devoicing process (Hualde
\& Colina, 2015). Nevertheless, in those modern varieties where the /s/
in coda position is produced, such as in Peninsular Spanish, this
voiceless phoneme might assimilate to the following sound, especially if
it is a voiced consonant (\emph{mismo}, {[}mízmo{]}), rather than a
vowel (\emph{las alas}, {[}lasálas{]}) (Hualde \& Colina, 2015; Hualde
et al., 2020). Hence, Spanish inventory does not have the phoneme /z/,
but {[}z{]} may occur as an allophone of /s/, due to a voice
assimilation process. It is important to mention, though, that it has
been claimed that this voicing is generally not complete, but partial
(Benet et al., 2012; Hualde \& Colina, 2015; Hualde et al., 2020).

There is some evidence of the influence of Spanish on the production of
the Catalan alveolar fricative, in contexts where, according to
prescriptivist rules (Institut d'Estudis Catalans, 2022), the {[}z{]}
should be produced, but {[}s{]} is produced instead. This phenomenon has
mainly been attested in those areas where there has been an extended
contact between Catalan and Spanish, such asthe so-called ``franja de
Aragón'' (borderland between Catalonia and Aragon), some areas of the
Valencian Community, and Barcelona and other urban areas (Benet et al.,
2012; Davidson, 2020; Recasens Vives, 1996, p.~268). The opposite
pattern has also been observed - that is, the voicing of the alveolar
fricative in the Spanish spoken in Barcelona and Valencia.
Interestingly, although in both Barcelona and Valencia Catalan and
Spanish coexist and influence each other, it seems that Catalan has a
greater influence on Spanish in Barcelona, but Spanish has a greater
influence on Catalan in Valencia (Davidson, 2020).

@@

The social meaning and the perception of intervocalic {[}z{]} in Spanish
has been studied in a variety of contexts, including both monolingual
(Chappell, 2016, 2017, 2021; Chappell et al., 2023; Chappell \& García,
2017; García, 2015, 2019; Silverstein, 2003) and bilingual contexts
(Davidson, 2014, 2015, 2019; Sinner, 2002), despite having been
categorized as a low-salient variant (Chappell, 2017, 2021; Chappell et
al., 2023; García, 2015)With respect to the former, it has been found
that for listeners from San José, Costa Rica, the covert attitudes
associated with intervocalic {[}z{]} vary significantly by speaker
gender. Consequently, male speakers who use {[}z{]} tend to be
positively evaluated, and generally perceived as nice, local, confident,
and masculine. Conversely, female speakers who use {[}z{]} are only
afforded negative associations of lower education and social class
(Chappell, 2016; Chappell et al., 2023; Silverstein, 2003).

Similarly, García (2019) shows that, in Ecuador, intervocalic /s/ is
evaluated differently in male and female speakers. Accordingly, women
who produce intervocalic {[}z{]} are viewed as significantly lower
status, less pleasant, and older. Contrary, men who produce intervocalic
{[}z{]} do not receive these negative evaluations. Additionally, this
variable has been found to serve as a marker of regional identity, with
voicing being associated with neighboring Highland Ecuadorian dialects
(Cuenca and Quito) that exhibit systematic voicing. García (2020)
hypothesizes that young men are the leaders of this change, essentially
due to the fact that they are in greater contact with these highland
varieties. Interestingly, the mean voicing rates of females are roughly
two age groups behind those of males. Consequently, it is believed that
it is not necessarily the case that women in Ecuador are avoiding
voicing like San José women due to its negative evaluations, but rather
that they are further behind in the adoption of the incoming variant.

With respect to the social meaning and the perception of intervocalic
{[}z{]} in bilingual contexts, this phenomenon has been mainly studied
in Catalan-Spanish contact situations, given its robust usage by
Catalan-Spanish bilinguals and, in particular, by those with greatest
exposure to and usage of Catalan (Davidson, 2014, 2015). According to
Sinner (2002), intervocalic {[}z{]} can be considered a possible
linguistic marker of Catalonian Spanish, despite its relative lack of
social awareness (Davidson, 2019; Sinner, 2002). In their study, Sinner
(2002) interviewed 12 speakers of Barcelonan Spanish and monolingual
Spanish-speakers from Madrid aged 27 to 41 regarding their awareness of
linguistic features of Catalanized Spanish. Results showed that only
select Barcelona speakers perceived intervocalic {[}z{]} as a linguistic
marker of Catalonian Spanish, whereas monolingual Spanish-speakers from
Madrid did not. Additionally, Davidson (2019) found that, in Barcelona,
the indexical field of intervocalic {[}z{]} features exclusively
positive associations with bilingualism and local solidarity, which
warrants its gradual adoption in the speech community, unlike other
phonetic markers of Catalonian Spanish that are the topic of overt
social ridicule, such as lateral velarization. These results are in line
with previous findings on intervocalic {[}z{]} in Catalonian Spanish.
For instance, Silverstein (2003) found that intervocalic {[}z{]} indexes
a local Barcelona identity, and the variant then acquires positive
associations with bilingualism and cultural pride. Overall, these
findings suggest that intervocalic {[}z{]} likely functions as a
sociolinguistic indicator of Catalonian Spanish within the Catalonian
speech community, indexing Catalan-Spanish bilingual membership while
lacking any considerable degree of generalized conscious awareness
(Davidson, 2019).

>>>>

Interestingly, Medieval Spanish, similarly to Modern Catalan,
distinguished between /s/ and /z/ in intervocalic position within a
word, although these sounds merged through a devoicing process
\cite{hualde_colina_2015}. Nevertheless, in those modern varieties where
the /s/ in coda position is produced, such as in Peninsular Spanish,
this voiceless phoneme might assimilate to the following sound,
especially if it is a voiced consonant (\emph{mismo}, {[}mízmo{]}),
rather than a vowel (\emph{las alas}, {[}lasálas{]}) \textbf{(Hualde \&
Colina, 2015; Hualde et al., 2020)}. Hence, Spanish inventory does not
have the phoneme /z/, but {[}z{]} may occur as an allophone of /s/, due
to a voice assimilation process. It is important to mention, though,
that it has been claimed that this voicing is generally not complete,
but partial \textbf{(Benet et al., 2012; Hualde \& Colina, 2015; Hualde
et al., 2020)}.

There is some evidence of the influence of Spanish on the production of
the Catalan alveolar fricative, in contexts where, according to
prescriptivist rules \textbf{(Institut d'Estudis Catalans, 2022)}, the
{[}z{]} should be produced, but {[}s{]} is produced instead. This
phenomenon has mainly been attested in those areas where there has been
an extended contact between Catalan and Spanish, such asthe so-called
``franja de Aragón'' (borderland between Catalonia and Aragon), some
areas of the Valencian Community, and Barcelona and other urban areas
\textbf{(Benet et al., 2012; Davidson, 2020; Recasens Vives, 1996,
p.~268)}. The opposite pattern has also been observed - that is, the
voicing of the alveolar fricative in the Spanish spoken in Barcelona and
Valencia. Interestingly, although in both Barcelona and Valencia Catalan
and Spanish coexist and influence each other, it seems that Catalan has
a greater influence on Spanish in Barcelona, but Spanish has a greater
influence on Catalan in Valencia \textbf{(Davidson, 2020)}.

@@

The social meaning and the perception of intervocalic {[}z{]} in Spanish
has been studied in a variety of contexts, including both monolingual
\textbf{(Chappell, 2016, 2017, 2021; Chappell et al., 2023; Chappell \&
García, 2017; García, 2015, 2019; Silverstein, 2003)} and bilingual
contexts \textbf{(Davidson, 2014, 2015, 2019; Sinner, 2002)}, despite
having been categorized as a low-salient variant \textbf{(Chappell,
2017, 2021; Chappell et al., 2023; García, 2015)}. With respect to the
former, it has been found that for listeners from San José, Costa Rica,
the covert attitudes associated with intervocalic {[}z{]} vary
significantly by speaker gender. Consequently, male speakers who use
{[}z{]} tend to be positively evaluated, and generally perceived as
nice, local, confident, and masculine. Conversely, female speakers who
use {[}z{]} are only afforded negative associations of lower education
and social class \textbf{(Chappell, 2016; Chappell et al., 2023;
Silverstein, 2003)}.

Similarly, \textbf{García (2019)} shows that, in Ecuador, intervocalic
/s/ is evaluated differently in male and female speakers. Accordingly,
women who produce intervocalic {[}z{]} are viewed as significantly lower
status, less pleasant, and older. Contrary, men who produce intervocalic
{[}z{]} do not receive these negative evaluations. Additionally, this
variable has been found to serve as a marker of regional identity, with
voicing being associated with neighboring Highland Ecuadorian dialects
(Cuenca and Quito) that exhibit systematic voicing. \textbf{García
(2020)} hypothesizes that young men are the leaders of this change,
essentially due to the fact that they are in greater contact with these
highland varieties. Interestingly, the mean voicing rates of females are
roughly two age groups behind those of males. Consequently, it is
believed that it is not necessarily the case that women in Ecuador are
avoiding voicing like San José women due to its negative evaluations,
but rather that they are further behind in the adoption of the incoming
variant.

With respect to the social meaning and the perception of intervocalic
{[}z{]} in bilingual contexts, this phenomenon has been mainly studied
in Catalan-Spanish contact situations, given its robust usage by
Catalan-Spanish bilinguals and, in particular, by those with greatest
exposure to and usage of Catalan \textbf{(Davidson, 2014, 2015)}.
According to \textbf{Sinner (2002)}, intervocalic {[}z{]} can be
considered a possible linguistic marker of Catalonian Spanish, despite
its relative lack of social awareness \textbf{(Davidson, 2019; Sinner,
2002)}. In their study, Sinner (2002) interviewed 12 speakers of
Barcelonan Spanish and monolingual Spanish-speakers from Madrid aged 27
to 41 regarding their awareness of linguistic features of Catalanized
Spanish. Results showed that only select Barcelona speakers perceived
intervocalic {[}z{]} as a linguistic marker of Catalonian Spanish,
whereas monolingual Spanish-speakers from Madrid did not. Additionally,
Davidson (2019) found that, in Barcelona, the indexical field of
intervocalic {[}z{]} features exclusively positive associations with
bilingualism and local solidarity, which warrants its gradual adoption
in the speech community, unlike other phonetic markers of Catalonian
Spanish that are the topic of overt social ridicule, such as lateral
velarization. These results are in line with previous findings on
intervocalic {[}z{]} in Catalonian Spanish. For instance,
\textbf{Silverstein (2003)} found that intervocalic {[}z{]} indexes a
local Barcelona identity, and the variant then acquires positive
associations with bilingualism and cultural pride. Overall, these
findings suggest that intervocalic {[}z{]} likely functions as a
sociolinguistic indicator of Catalonian Spanish within the Catalonian
speech community, indexing Catalan-Spanish bilingual membership while
lacking any considerable degree of generalized conscious awareness
\textbf{(Davidson, 2019)}.

\section{Methods}

\subsection{Materials}

Data were collected using an online survey hosted on Qualtrics that
consisted of two parts. The first part, a demographic questionnaire,
collected information regarding participant age, gender, place of
origin, place of residence, mother tongue, family income, and level of
studies. The second portion of the survey consisted of a perception task
where participants were presented with a series of stimuli containing
the target phoneme contrast (intervocalic voiced and voiceless alveolar
fricative) followed by 11 evaluative statements related to social
characteristics such as level of studies, friendliness, social class,
and ethnic origin. These evaluative statements were then followed by a
100-point draggable sliding scale with which participants indicated
their level of agreement or disagreement with the content of the
statement, where 0 was ``totally disagree'' and 100 was ``totally
agree''.

The target stimuli were recorded by a 28-year-old female native speaker
of Catalan in a whisper room 6084 E sound booth using a personal
communication device. The stimuli were between 3 and 4 seconds long and
were controlled for utterance length (5 words), syllable count (2
syllables for the target word and 3 for the following adjective), and
lexical stress (target sound in unstressed position). A total of 12
target stimuli containing intervocalic voiced /s/ were recorded, as well
as different samples of the voiceless /s/. The target sentences were
manipulated in Praat by splicing the voiced /s/ and pasting the
voiceless /s/ in order to create 2 conditions: 12 stimuli containing
intervocalic voiced /s/ and 12 stimuli containing intervocalic voiceless
/s/, for a total of 24 experimental sentences. These sentences were
combined with 32 fillers targeting other phonemic contrasts, for a total
of 56 sentences randomly distributed in two different versions. Each
included 12 experimental stimuli (6 x 2 conditions) and 16 fillers (8 x
2 conditions).

\subsection{Procedure}

Participants self-identified for this study. Recruitment materials in
Catalan with information about the study, the requirements to
participate and a link to the survey were posted on different social
networks and online communication channels, such as WhatsApp and
Instagram. Participants who were interested and wanted to participate
clicked on the link and read the study description. Those who decided to
participate completed the survey on their own electronic devices and at
their own pace without being observed. On average, the survey took
approximately 15 minutes to be completed. The survey followed a
between-subjects design. Therefore, participants were randomly assigned
to one of the two versions.

\subsection{Participants}

A total of 78 responses to the questionnaire were collected. Of these,
56 responses were incomplete and were therefore discarded. Data are
drawn from a total of 22 participants (13 females) aged between 18 and
72 years old (mean = 38) from Barcelona and Girona. In order to be
eligible to take part in the study, participants had to meet the
following three requirements: to be over 18 years of age, to be from
Catalonia, and to be proficient in both Catalan and Spanish.

\subsection{Data analysis}

Data were analyzed in R (R Core Team, 2023). First, we performed a
descriptive analysis that consisted of creating a series of plots and
tables to summarize and describe the features of the dataset and to
identify potential trends. After that, we performed inferential
statistical analysis. For each of the 11 evaluative statements, we
created a series of hierarchically nested linear regression models using
the lme4 package (Bates et al., 2015) with attitudinal sliding scale
responses as the dependent variable, and phoneme, mother tongue, and
province as potential predictors. To determine the best fit model for
each statement, we used the anova() function (Fox \& Weisberg, 2019).
